The Regional Effects of Marine Protected Areas

### 2020-03-26 documents/ovando-regional-effects-of-mpas-v2.tex
passing to squid since sf on mac is broken beyond belief

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: danovan@uw.edu}

@@

Indeed, many recent agreements to expand MPAs (most proximately, the
Convention on Biological Diversity's Strategic Plan for Biodiversity,
which calls for 10\% of coastal waters to be protected inside MPAs by
2020 and the International Union for Conservation of Nature call for
30\% by 2030) rely on the belief that well-designed MPAs will achieve
benefits both within, and outside, their borders. Despite these
assumptions, our collective scientific understanding of the
regional-scale conservation and fishery impacts of current and future
MPAs is surprisingly limited.

What is the scientific evidence of the conservation effects of MPAs?
Numerous studies provide evidence that well-enforced and appropriately
sized MPAs can produce conservation benefits within their borders
(5--8). As these conservation benefits accrue inside MPAs, theory holds
that MPAs can affect the waters beyond their borders through the
spillover of adult and larval fish from the protected to the fished
areas, as well as through displacement of fishing effort. Several
studies have documented empirical evidence for the existence of adult or
larval spillover affecting both abundance and fisheries (9--17), as well
as alteration of fishing effort in reaction to (18--20) and in
anticipation of (21) MPA placement. The potentially more important
question, however, is not whether spillover occurs (it must to some
degree in any realistic scenario), but what the net effects of spillover
are and whether those effects are empirically detectable. From a fishery
perspective, are spillover benefits sufficient to offset losses in
fishing grounds and changes in responses of displaced fishers caused by
an MPA? From a conservation perspective, how much does the buildup of
fish inside an MPA increase biomass outside the protected area? Overall,
what are the regional effects of MPAs?

As stakeholders around the world increasingly seek to use MPAs in marine
resource management portfolios, and the performance of existing MPAs is
evaluated, it is critical that we develop a better understanding of the
magnitude and drivers of regional-scale MPA effects. To address this
gap, this study examines two critical questions: 1) What do we expect
the regional-scale conservation effects of MPAs to be and 2) When (and
how) can we expect to empirically detect these effects? We answer these
questions using a simulation analysis to frame the theoretical regional
conservation impacts of MPAs. We then present an empirical test of the
evidence for regional-level conservation effects of MPAs resulting from
a network of closures put in place in the Channel Islands, California,
USA, in 2003.

@@

Much of the literature on the effects of MPAs focuses on assessing the
conservation effects within the borders of protected areas (7). While
these within-MPA effects are vitally important for protecting rare
species, biodiversity, critical habitats, and often tourism, they paint
an incomplete picture of the overall population effects of MPAs. The
organisms within the borders of protected areas are generally part of a
broader biological stock, connected through adult movement and larval
dispersal. If the goal of conservationists or natural resource managers
is to increase the total abundance or productivity of a resource, a
broader question we should ask of MPAs is not just are there more fish
inside their borders, but also how have the reserves affected abundances
throughout the region in which they are located? This logic, that MPAs
will have conservation benefits for most species beyond their borders,
is implicit in all multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in
total biomass of fish (summing inside and outside of MPAs) relative to
the total biomass of fish that would have occurred without the MPAs
(acknowledging that other outcomes such as increased biodiversity or
resiliency are also important to conservation but are beyond the scope
of this analysis).

We used a simulation model to examine the regional conservation effects
of MPAs under a wide array of conditions that might reasonably be
expected to affect these outcomes. The model explores combinations of
MPA designs, life histories, and fishing fleet dynamics (See Table.S1
for a complete description of simulation variables). We use this model
to explore how different drivers interact to affect the conservation
outcomes of MPAs.

\hypertarget{conservation}{%
\subsubsection*{Theoretical Conservation Effects}\label{conservation}}
\addcontentsline{toc}{subsubsection}{Theoretical Conservation Effects}

As a thought experiment, imagine a region that has driven its fish
populations to near extinction and then places 100\% of its waters
inside a no-take MPA. In that setting, we would expect the
regional-scale conservation benefits to be massive. On the other hand,
imagine implementing an MPA in an area where a sedentary species has
been only lightly fished. In that setting, it is possible to create a
small net conservation loss if the concentration of fishing pressure
outside the reserve has a greater effect than the biomass buildup inside
the reserve. Within these broad bounds, numerous factors can act to
affect the regional effects of MPAs. These include the scale of adult
and larval dispersal relative to the size of the MPAs (22--25), the
strength and timing of density dependence in the population (e.g.~pre-
or post-settlement), how overfished the population would be without the
MPA, and how fishing activity and management responds to the
implementation of the MPAs (4, 8, 26--32). In addition, even for the
same total area of MPAs, the location and spacing of the MPAs can have a
profound influence on their cumulative impact through habitat and
network effects (4, 33).

We tested the effects of different combinations of these theoretical
drivers of MPA conservation effects using over 10,000 simulated MPA
scenarios. A striking outcome of this simulation analysis is how the
level of fishing pressure can cause dramatic and predictable shifts in
conservation outcomes. We measure fishing pressure by the degree of
simulated depletion (percent of unfished biomass gone, such that a
depletion of 0\% means the population is at 100\% of unfished biomass,
100\% means 0\% of unfished biomass remains) in the scenario without
MPAs. When MPAs were placed in relatively unexploited fisheries
(depletions less than 40\%), the median regional conservation effect was
barely noticeable (4\%). For ecosystems fished near MSY levels
(depletion of 40\%-60\%), the median conservation effect was 28\%, while
for overfished ecosystems (without-MPA depletions above 60\%) the median
regional conservation effect was over 100\%
(Fig.\ref{conservation-effects}-A). Across all of these groups though,
effects as low as declines of -50\% and as high as an increase of more
than 100\% were possible. It should be noted though that these runs
represent the range of scenarios evaluated in our simulation, and that
certain types of scenarios may be much more likely than others in the
real world. To put these results into context, the FAO estimates that
7\% of the worlds fisheries with status estimates fall into the
relatively unexploited category, 60\% fall into the fully fished
category, and 33\% fall into the heavily fished category (34), though
works that include a broader range of fisheries estimate that 50\% or
more of stocks to fall into the heavily fished category (35, 36).

These simple results support the intuitive conclusion that bigger MPAs
on overfished populations produce large positive conservation outcomes.
However, when combined with other factors such as fleet dynamics, life
history, and MPA size, the conservation effects of MPAs can vary widely
in magnitude and timing, a finding corroborated by empirical evidence
(37--40) (Fig.\ref{conservation-effects}-B). As the area protected
inside MPAs increases, positive effects become more likely, but even
MPAs protecting most of a population's footprint can produce
conservation effects as low as 0\% and as high as 200\%. The degree of
depletion (i.e.~fishing pressure) has a clearer signal, with small
conservation benefits of MPAs when depletion is low, and larger effects
when depletion is high, but again even for severely depleted populations
the conservation effects of MPAs can vary widely (although nearly always
with population benefits).

While positive conservation outcomes were much more likely to occur
across our simulations (94\% of simulations), the 6\% of runs that
produced net conservation losses are of particular interest. These runs
occurred almost exclusively when the simulated fleet followed a
``constant-catch'' fishing strategy. Under the constant-catch fleet
model, fishing communities (or fishery policies) seek to catch the same
amount regardless of the presence of an MPA. While a constant-catch
greater than MSY is not possible over the long-term under the
assumptions of our model, over the short-term a constant-catch scenario
is not implausible. Subsistence fisheries may use a constant-catch style
policy over the short-term, as they seek to ensure that their food needs
are met. More industrial fisheries may have pre-arranged agreements with
buyers to deliver set amounts of fish. Constant-catch dynamics might
also occur in fisheries with constraining quotas that are not updated
after the implementation of MPAs. These patterns suggest that coupling
MPA implementation with appropriate management reforms that reduce the
effects of constant-catch strategies can move the predicted conservation
effects of MPAs to being nearly uniformly positive. The possibility of
constant-catch strategies to result in net MPA conservation losses
highlights the critical importance of understanding the broader
management context in which MPAs are placed.

\begin{figure*}%[tbhp]
  \centering
  \includegraphics[width=.9\linewidth]{figs/conservation-effect.png}
  \caption{Median (A) and range (B) of equilibrium regional MPA conservation effects (percent change in total biomass with MPAs relative to without MPAs) across a range of depletion and MPA sizes (and incorporating the full range of scenarios included in our study). "Range in MPA" is the percent of patches covered by an MPA, "Depletion" is the depletion that would have occurred in equilibrium without the MPA. Colors in panel B represent the percentage of simulations that fell into an effect size bin at each x-axis bin, such that the sum of all percentages for a given x-axis bin is 1.}
  \label{conservation-effects}
\end{figure*}

@@

How can we empirically measure the regional effects of MPAs? Intuition
and bio-economic theory tells us that the regional conservation effects
of MPAs are likely to be highly context dependent across locations and
across species within locations, so careful thought must be devoted to
an empirical design to detect their effects. The perfect experiment
would involve two parallel worlds that were identical, except for the
implementation of an MPA. In world ``A'', no MPA would be implemented,
and in the facsimile world ``B'', the MPA would be implemented. Both
worlds would be tracked before and after MPA implementation, and the
conservation and fishery outcomes would be compared after treatment.
Instead of two parallel worlds, a similar experimental design would
involve random placement of MPAs.

Unfortunately, neither of these experimental designs has been
implemented, because parallel worlds do not exist and MPAs are never
placed at random. Careful causal inference requires some other means of
controlling for biases introduced by factors such as the MPA siting
process and biological spillover and concentration of fishing effort
(41, 42).

Conservation effects are typically estimated by comparing organism
densities or biomass inside MPAs to densities or biomass in selected
control sites outside MPAs, which we will refer to as response ratios.
(5) and (43) present meta-analyses of hundreds of such studies. These
results often find massively higher densities and biomasses inside MPAs
than outside (5). While response ratios are intended as a measure of the
effectiveness of reserves within their borders, these result stand in
contrast to the smaller region wide effect sizes predicted by our
simulation model. This raises the question: How reliable are density or
biomass ratios as estimators of regional conservation effects?

Control sites used in density or biomass ratios are often selected based
on abiotic or ecological traits such as habitat characteristics (41).
However, selection of control sites is further complicated by the very
spillover that MPAs are often intended to create. Export of adults or
larvae from the MPA to the ``control'' site affects their status as
controls, as does displacement of fishing effort from MPAs to control
sites. In theory, control sites far enough away to negate both
biological spillover and concentration by the fishing fleet could be
selected, but finding suitably far sites that are also appropriate
proxies for the ecological and economic context of the MPAs may be
challenging. While these concerns have been stated previously (44), the
MPA evaluation literature has by and large been unable to adequately
address them, often due to the very real challenges of acquiring the
necessary data (41).

Failure to account for these spillover effects can result in a biased
estimate of the true effect of a policy such as MPA placement (42). In
an ideal setting, the control site used in a response ratio would be a
perfect counterfactual for what would have occurred without the MPA. We
used our simulation model to approximate this scenario, calculating the
response ratio as the ratio of densities inside the MPAs relative to the
overall density from the paired simulation without MPAs multiplied by
the area of the population in an MPA (to account for the fact that we
would expect the effectiveness of a response ratio as a proxy for
regional conservation effects to scale with the size of the MPA relative
to the range of the species, Fig.\ref{density-ratio}-A). In this
idealized case, the response ratio is often a reasonable estimator for
the regional conservation effect.

What happens when the ``control'' sites are in fact affected by the
application of the MPA through biological spillover and fishing
concentration? In order to address this possibility, we calculated a new
response ratio as the mean density inside MPAs relative to the mean
density outside the MPAs (both weighted by distance from MPA borders and
scaled by MPA size). We only include simulations in which habitat and
larval dispersal rates are identical inside and outside of MPAs, to
approximate a scenario in which treatment and control sites have been
paired by ecological characteristics. Under these circumstances the
response ratio is a somewhat biased and very inaccurate estimator of the
true regional effect of MPAs.

While on average high response ratios correspond with high true regional
effects, the variance is high, so the response ratio alone cannot be
used as a reliable proxy for the true regional effect, without
accounting for the sources of biases discussed here. For higher movement
rates, response ratios were frequently near 0\% (potentially leading
stakeholders to conclude that the MPA had been ineffective), when in
fact in many cases the true effect of the MPA was highly positive. For
highly sedentary species, extremely high response ratios could create
the appearance of massive conservation gains when in fact the net effect
on the regional population has been near zero or even negative
(Fig.\ref{density-ratio}-B).

We note that spatial before-after-control-impact (BACI) studies present
a potential improvement over response ratios, controlling for
differences in mean densities in the treated and control regions as well
as MPA independent trends, but are much rarer due to the need for
extensive pre-MPA monitoring. However, spatial BACI still requires that
the MPAs do not affect the control sites, and as such has many of the
same limitations as response ratios as an estimator of regional
conservation effects.

\begin{figure*}%[tbhp]
  \centering
  \includegraphics[width=.9\linewidth]{figs/density-ratio.png}
  \caption{Simulated MPA-size weighted response ratios (x-axis) plotted against true regional change in biomass caused by MPAs (y-axis). Each point represents a simulation, and color represents the movement rate of fish in the simulation (expressed as a percentage of the patches a fish is likely to move in one time step). The black line shows the one:one fit line, the blue line is a linear fit of the relationship between observed response ratios and true regional change in biomass. Inside vs. Experiment simulates a true control, while Inside Vs. Outside uses a control that may be affected by the treatment.}
  \label{density-ratio}
\end{figure*}

\hypertarget{estimating}{%
\subsection*{MPA Effects in the Channel Islands}\label{estimating}}
\addcontentsline{toc}{subsection}{MPA Effects in the Channel Islands}

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/channel-islands.pdf}
  \caption{Map of PISCO sampling locations within the the Channel Islands National Marine Sanctuary, California, USA. Grey boxes indicated MPAs, points sampling locations, with color representing whether the site is inside an MPA (blue) or not (red)}
  \label{channel-islands}
\end{figure}

Given these challenges, what is a viable alternative for estimating the
conservation effects of MPAs? In a manner similar to (38), we use a
difference-in-difference estimator comparing densities in finfish
species targeted by fishing effort (i.e., those potentially affected by
an MPA) from those not targeted before and after MPA implementation
(42). Targeted species in the Northern Channel Islands, California,
include commercially important finfish such as California sheephead
(\emph{Semicossyphus pulcher}), and copper (\emph{Sebastes caurinus})
and blue (\emph{Sebastes mystinus}) rockfish. Each of these targeted
species was the subject of prior bio-economic modeling related to the
effects of MPAs in southern California (45, 46). But, we must note that
the analysis omits species from important invertebrate fisheries
including red urchin (\emph{Mesocentrotus franciscanus}) and spiny
lobster (\emph{Panulirus interruptus}). Non-targeted species include
garibaldi (\emph{Hypsypops rubicundus}), halfmoons (\emph{Medialuna
californiensis}), and blacksmith (\emph{Chromis punctipinnis}) (See
Table.S2 for a complete list of species). The key assumptions of this
approach are that a) within the time-frame of the model there are no
significant interaction effects between the targeted and non-targeted
species (which in fact we do not detect, see SI), and b) that in the
absence of the MPAs both the targeted and non-targeted groups of species
would have exhibited similar trends in densities. The advantage to this
approach is that given the time-frame of the model (16 years), we
believe that spillover effects of MPA placement on the non-targeted
species are likely to be much less severe than the effects of biological
spillover and fishing concentration that may bias the performance of
estimators such as a response ratios or a spatial BACI design.

We applied this difference-in-difference strategy to a large dataset in
California. We used empirical kelp forest survey data from the
Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO)
monitoring in the Northern Channel Islands with the ultimate goal of
testing the regional effects of MPAs in a real world context. A network
of MPAs covering approximately 20\% of the islands' waters was put in
place in 2003 as part of the California Marine Life Protection Act
(MLPA) (see (47), (48), (49), and (50) for information on the creation
of the MLPA). PISCO conducts visual SCUBA surveys at a large number of
rocky-reef and kelp forest sites inside and outside of MPAs throughout
the Channel Islands, producing estimates of densities of fishes that are
both targeted and non-targeted by fishing
(Fig.\ref{channel-islands}-\ref{pop-trends}).

Although the Islands do not encompass the full population of any of
these species, they encompass a large enough area to assess the regional
MPA conservation effects we might see in the Channel Islands after 13
years of protection (the time span covered by our model). The median
regional conservation effect produced by our simulation model after 13
years of protection by MPAs covering between 15\%-20\% of a population's
range was 25\%. This projected effect has an an interquartile range of
8\% and 100\%, although values as high as a 200\% increase and as low as
a 50\% decrease were also produced. Our expectation then is that all
else being equal we are likely looking for a modest but positive effect
size.

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/pop-trends.pdf}
  \caption{Centered and scaled mean annual biomass densities of included targeted (red) and non-targeted (blue) fish species (faded lines) and smoothed means with 95\% confidence intervals over time (darker lines and ribbons). Vertical red dashed line shows year of MPA implementation.}
  \label{pop-trends}
\end{figure}

The raw trends in the biomass densities appear to show parallel trends
in the densities of targeted and non-targeted fishes pre-MPA, with the
targeted group diverging first positively and then negatively over time
(Fig.\ref{pop-trends}). However, these raw trends could be caused by any
number of confounding variables besides the effects of the MPAs
themselves, from changes in observer skill and sampling locations, to
shifts in environmental conditions. Using our difference-in-difference
model in an effort to control for these factors, we found no significant
difference in the biomass densities of targeted and non-targeted fishes
pre-MPA (before 2003). Following the implementation of the MPAs in 2003,
we see evidence of an increasing trend in densities of targeted relative
to non-targeted fish, reaching a peak value in 2014. After 2014,
densities of targeted species appear to have declined relative to what
we would expect based on the non-targeted group (Fig.\ref{did-plot}).

\begin{figure}%[tbhp]
  \centering
  \includegraphics[width=1\linewidth]{figs/did_plot.png}
  \caption{Estimated divergence in biomass densities of targeted and non-targeted fishes throughout the Channel Islands (integrated across inside and outside of MPAs). MPAs are implemented in 2003 (red dashed line). Colored cells in the background indicate the frequency of MPA conservation effects predicted by our simulation modeling for MPAs similar in design and species composition to the Channel Islands in each year.}
  \label{did-plot}
\end{figure}

@@

MPAs are an important part of the marine resource management toolbox.
Under ideal circumstances they can protect individual species and
ecosystem linkages, while supporting local economies through tourism and
fishing opportunities. One rationale for the expansion of MPAs is that
they will deliver net conservation benefits both inside \emph{and
outside} their borders. To date, this assumption is insufficiently
tested, and this is the focus of our paper. Our results show that
regional conservation benefits of MPAs are highly context dependent and
in many circumstances, are likely to be so small that they are nearly
impossible to detect empirically. Indeed, this is exactly what we found
in our empirical case study from the Channel Islands, California, USA.

We find regional conservation gains from MPAs in 94\% of our simulations
(Fig.\ref{conservation-effects}). MPAs covering less than 25\% of the
population range were unlikely to produce regional conservation gains
above 10\% unless the stock would have been severely overfished without
the MPA. Given that few marine species on the planet have more than 25\%
of their entire range protected in MPAs, this suggests that large
regional gains are unlikely to be common unless the overall extent of
MPAs grows globally. Large MPAs protecting highly depleted stocks almost
always produced large regional conservation gains, as should be
expected.

The median conservation effect of 36\% stands in contrast to the large
within-MPA effects on biomass densities reported by (51) and (5). This
difference is due in part to the fact that within-MPA effects are likely
to be much larger than regional-MPA effects, but may also reflect the
many challenges of translating response ratios into regional
conservation effects. Our simulation analysis shows that without proper
controls (in particular for the effects of fleet redistribution),
response ratios well over 100\% are entirely plausible even when the
actual conservation effect is much lower (and \emph{vice versa}).

>>>>

\emph{as we move away from ``set it and forget it'' management towards
adaptive management of marine ecosystems, it is critically important
that we develop robust strategies for monitoring the effectiveness of
our management intercentions}

@@

Indeed, many recent agreements to expand MPAs (most proximately, the
Convention on Biological Diversity's Strategic Plan for Biodiversity,
which calls for 10\% of coastal waters to be protected inside MPAs by
2020 and the International Union for Conservation of Nature call for
30\% by 2030) are based on the expectaion that well-designed MPAs will
achieve benefits both within, and outside, their borders. Despite these
assumptions, our collective scientific understanding of the
regional-scale conservation and fishery impacts of current and future
MPAs is surprisingly limited.

Numerous studies provide evidence that well-enforced and appropriately
sized MPAs can produce conservation benefits within their borders
(5--8). As these conservation benefits accrue inside MPAs, theory holds
that MPAs can affect the waters beyond their borders through the
spillover of adult and larval fish from the protected to the fished
areas, as well as through displacement of fishing effort. Several
studies have documented empirical evidence for the existence of adult or
larval spillover affecting both abundance and fisheries (9--17), as well
as alteration of fishing effort in reaction to (18--20) and in
anticipation of (21) MPA placement. The potentially more important
question, however, is not whether spillover occurs (it must to some
degree in any realistic scenario), but what the net effects of spillover
are and whether those effects are empirically detectable. From a fishery
perspective, are spillover benefits sufficient to offset losses in
fishing grounds and changes in responses of displaced fishers caused by
an MPA? From a conservation perspective, how much does the buildup of
fish inside an MPA increase biomass outside the protected area? Overall,
what are the regional effects of MPAs?

As stakeholders around the world increasingly seek to use MPAs in marine
resource management portfolios, it is critical that we develop a better
understanding of the magnitude and drivers of regional-scale MPA
effects. This is particuarly true in the context of a changing climate,
in which ``set it and forget it'' management strategies are increasinly
untenable; we need to be able to monitor the performance of MPAs in
order to adapt them as dictated by a shifting environment. The Channel
Islands National Marine Sanctuary, California, USA provides an ideal
case study to address this need. A network of protected areas covering
approximately 20\% of the Islands' waters was put in place in 2003 (22).
This MPA network has subsequently been used as a model case study in
protected area design around the world XX.

We use data from the first 14 years of protection to provide what is to
our knowledge the first empirical assessment of the regional effect of a
large MPA network on a wide array of finfish species. In contrast to
clear differences in biomass densities observed inside and outside of
well protected MPAs both globally (5) and in the Channel Islands
({\textbf{???}}) we are unable to detect a clear regional effect from
the Channel Islands MPAs. We build off of existing MPA theory to
consider why this might be, and provide guidance for scientists and
managers as to when and how we might expect to estimate the regional
conservation effects of MPAs.

@@

The empirical MPA literature has generally focused on assessing
conservation effects within the borders of protected areas (7). While
these within-MPA effects are vitally important for protecting rare
species, biodiversity, critical habitats, and often tourism, they paint
an incomplete picture of the overall population effects of MPAs. The
organisms within the borders of protected areas are generally part of a
broader biological stock, connected through adult movement and larval
dispersal. If the goal of conservationists or natural resource managers
is to increase the total abundance or productivity of a resource, a
broader question we should ask of MPAs is not just are there more fish
inside their borders, but also how have reserves affected populations
throughout the region in which they are located? This logic, that MPAs
will have conservation benefits for most species beyond their borders,
is implicit in all multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in
total biomass densities of fish (averaging inside and outside of MPAs)
relative to the total biomass densities of fish that would have occurred
without the MPAs (acknowledging that other outcomes such as increased
biodiversity or resiliency are also important to conservation but are
beyond the scope of this analysis). Numerous factors can affect the
regional effects of MPAs. These include the scale of adult and larval
dispersal relative to the size of the MPAs (8, 26--29), larval dispersal
patterns and the strength and timing of density dependence in the
population (e.g.~pre- or post-settlement, \textbf{???}), how overfished
the population would be without the MPA, and how fishing and management
responds to the implementation of the MPAs ({\textbf{???}}, 4, 30--39).
In addition, even for the same total area of MPAs, the location and
spacing of the MPAs can have a profound influence on their cumulative
impact through habitat and network effects (4, 40).

Given this existing body of theory, what should we expect \emph{a
priori} the effect size of the Channel Islands MPAs to be? While we know
many parameters will affect the expected effects of the MPAs, theory
suggests that life history, adult and larval dispersal relative to MPA
size, and fishing pressure will be key drivers. The targeted species in
our database span a range of life histories, but are largely made up of
fishes in the genus XX (e.g.) and \textbf{Sebastes} (e.g.~), with a
median Von Bertalanffy growth coefficient of XX, and a median age at
maturity of XX (41). The MPAs in this study cover 21\% of the the
surface waters of the Northern Channel Islands, a region spanning
roughly 90km east to west. While detailed dispersal studies are not
available for all of the species covered by our study, what information
we have suggest that while adults of some of these species are likely to
exhibit some site fidelity (42), larvae are likely distributed beyond
the northern Channel Islands (43). Therefore, if we assume that the
population of the targeted species is at least equal to the extent of
the Northern Channel Islands, we can assume that at most the MPAs cover
roughly 20\% of the targeted populations' ranges.

Turning to the critical question of fishing pressure, formal stock
assessments are largely lacking for these species. However, what
evidence we have suggests mostly moderate fishing pressure, with some
species such as California sheephead (\emph{Semicossyphus pulcher}) and
blue rockfish (\emph{Sebastes mystinus}) experiencing high levels of
fishing mortality / biomass levels below target levels during the early
2000s (44, 45). We integrated these network design features, life
history traits, and exploitation histories into a spatially explicit
bio-economic simulation model, and used this model to generate expected
outcomes for the Channel Islands MPAs. Our results suggest that while a
wide range of outcomes are plausible, from 0\% to upwards of 200\%,
across the majority of our simulations the effect size was 25\% or less,
with a median simulated effect size of roughly 20\%; a potentially small
value given the natural variability and sampling challenges of marine
environments.

@@

Having established a theoretical grounding for the likely biomass
density outcomes of the Channel Island MPAs, what empirical evidence do
we have of the effects? \textbf{???} examined changes in biomass
densities of species targeted by fishing activity inside and outside of
the MPAs over time (a metric generally termed a ``response ratio''), and
compared these changes to the trends in non-targeted species. They found
a statistically significant increase in the response ratios of targeted
species over time, and evidence that this increase is smaller in the
non-targeted species. Updating the results of (46) through 2017 with a
Bayesian response-ratio regression shows a continuation of the
increasing trend in the response ratios of targeted species
(Fig.@ref(fig:XX)).

Why do these response ratio results not serve as sufficient evidence for
the regional effects of the Channel Island MPAs? Control sites used in
calculating response ratios are often selected based on abiotic or
ecological traits such as habitat characteristics (47). However,
selection of control sites is further complicated by the very spillover
that MPAs are often intended to create. Export of adults or larvae from
the MPA to the ``control'' site affects their status as controls, as
does displacement of fishing effort from MPAs to control sites. In
theory, control sites far enough away to negate both biological
spillover and concentration by the fishing fleet could be selected, but
finding suitably far sites that are also appropriate proxies for the
ecological and economic context of the MPAs may be challenging. While
these concerns have been stated previously (48), the MPA evaluation
literature has by and large been unable to adequately address them,
often due to the very real logistical challenges of identifying and
sampling adequate control sites that are truly independent of the MPAs
in question. As a result of these challenges, response ratios can be a
highly imprecise and biased measure of the regional conservation effect
of an MPA network (47).

In the case of the Channel Islands MPAs, control sites are often located
within only a few kilometers of an MPA, suggesting that they are
susceptible to biological spillover and concentration of fishing effort
resulting from the MPAs. Given these complications, we can certainly
interpret the response ratios reported in (46) and updated here as
evidence that the MPAs are providing effective protection of targeted
fish biomass inside their borders, particularly in light of the general
increase in response ratios over time. However, it is unclear how well
these response ratio results serve as in indicator of overall regional
effects of the MPAs.

To illustrate this problem, we used a bio-economic model to simulate
response ratio trajectories for species and MPA coverage representative
of the northern Channel Islands (see methods and SI for details of model
structure). These simulations cover the life-histories and MPA sizes
seen in the Channel Islands, but vary in key unknowns such as the actual
degree of fishing pressure, the timing of density dependence, and the
fleet responses to the MPAs. For each year of protection, we paired our
simulated response ratios to the estimated posterior probability
distributions of the Channel Island response ratios. For example, if the
mean estimated response ratio in the year 2006 is one, we found all
simulations with three years of protection that had simulated response
ratios near one, and then pulled out the ``true'' simulated regional MPA
effect from each of those simulations. This provides us with a
distribution of simulated regional MPA effects that could plausibly
generate the types of response ratios actually observed in the Channel
Islands.

As reported in (46), we see evidence for an increasing trend in response
ratios of targeted species over time (Fig.@ref(fig:rr-plot)). This
provides strong evidence that the MPAs are providing protection for
targeted species within their borders. However, our simulation results
show that the response ratio trends we observe in the data could
plausibly be produced by a wide range of regional MPA effects
(Fig.@ref(fig:rr-plot)). Response ratios well over one were associated
with regional MPA effects generally less than 25\%, and many simulations
produced large response ratios but regional MPA effects close to 0\%
This can occur if for example fishing pressure is only moderate, adult
movement is low, larval dispersal is high, and displaced fishing effort
concentrates around the border of the MPAs.

\textbackslash begin\{figure\emph{\}\%{[}tbhp{]} \centering
\includegraphics[width=.9\linewidth]{figs/response_ratio_plot.png}
\textbackslash caption\{90\% Posterior probability distributions of
response ratios for targeted species (x-axis) over time (y-axis) in
blue. Simulated MPA effects associated with simulated response ratios
matched to empirical response ratios in red. For response ratios, a
value of zero indicates that biomass densities of targeted and
non-targeted species are identical inside and outside MPAs, a value of 1
indicates that biomass densities of targeted species are 100\% greater
inside MPAs than outside. For MPA effect, a value of zero indicates that
mean biomass densities are identical in the with- and without- MPA
scenarios. A value of 1 indicates that mean biomass densities are 100\%
greater in the scenario with MPAs than the scenario without MPAs.\}
\label{rr-plot} \textbackslash end\{figure}\}

Given the potential unreliability of response ratios for the task, in an
ideal world how would we empirically measure the regional effects of
MPAs? The perfect experiment would involve two parallel worlds that were
identical, except for the implementation of an MPA. In world ``A'', no
MPA would be implemented, and in the facsimile world ``B'', the MPA
would be implemented. Both worlds would be tracked before and after MPA
implementation, and the biomass densities would be compared after
treatment. Instead of two parallel worlds, a similar experimental design
would involve random placement of MPAs. Unfortunately, neither of these
experimental designs has been implemented, because parallel worlds do
not exist and MPAs are to our knowledge never placed at random. Careful
causal inference requires some other means of controlling for biases
introduced by factors such as the unobserved environmental shocks, the
MPA siting process, biological spillover, and concentration of fishing
effort (47, 49).

Building off of the concepts explored in (46), we propose an alternative
identification strategy utilizing biomass densities of XX species that
are not directly targeted by fishing as our control group
(non-targeted), and biomass densities of XX species targeted by fishing
as our treatment group. Targeted species in the Northern Channel
Islands, California, include commercially important fin-fish such as
California sheephead (\emph{Semicossyphus pulcher}), and copper
(\emph{Sebastes caurinus}) and blue (\emph{Sebastes mystinus}) rockfish.
Each of these targeted species was the subject of prior bio-economic
modeling related to the effects of MPAs in southern California (50, 51).
It should be noted though that our analysis omits species from important
invertebrate fisheries including red urchin (\emph{Mesocentrotus
franciscanus}) and spiny lobster (\emph{Panulirus interruptus}).
Included non-targeted species include garibaldi (\emph{Hypsypops
rubicundus}), halfmoons (\emph{Medialuna californiensis}), and
blacksmith (\emph{Chromis punctipinnis}) (See Table.S2 for a complete
list of species). We used a Bayesian difference-in-difference regression
to estimate any difference in mean total biomass densities of fin-fish
species targeted by fishing effort (i.e., those potentially affected by
an MPA) and those species not targeted by fishing before and after MPA
implementation (49). The result of this regression is, conditional on
the assumptions of the model, an estimate of the effect of the MPAs on
the mean total biomass densities of targeted species throughout the
Channel Islands.

As conventional wisdom and theory would suggest, over the first three
years of implementation (2003-2006), the effects of the MPAs are
unclear, with support for a small negative effect to a substantially
positive effect (with much higher probability of a small positive
effect). Over the next six years the model estimates greater
probabilities of an increasingly positive MPA effect, peaking in
2009-2011 with a median estimate of MPA effect of a xx increase in mean
total biomass of targeted species (90\% credible interval 37\% - 128\%.
These empirical estimates are in line with (though in the upper range
of) the outcomes that our simulation model suggests are plausible given
the rough characteristics of the northern Channel Island MPAs. However,
in the subsequent years the trend reverses itself, and for the years
2015-2017 we once again see no clear effect of the MPAs
Fig.@ref(fig:mpa-effect)

\textbackslash begin\{figure\emph{\}\%{[}tbhp{]} \centering
\includegraphics[width=.9\linewidth]{/Users/danovan/projects/regional-effects-of-mpas/documents/figs/mpa_effect_plot.png}
\textbackslash caption\{Results of difference-in-difference regression
estimating the regional effect of the northern Channel Island MPAs on
mean total biomass densities of targeted species (difference in mean
total biomass density of targeted species over time relative to expected
levels using non-targeted species as a control). Grey distributions show
90\%XX posterior probability distribution of estimated MPA effect; red
point is median estimated effect, thicker red section 50\% credible
interval, thinner red line 90\% credible interval). Blue distributions
in background show range of MPA effects produced by simulation model
tuned to reflect the dynamics of the Northern Channel Island MPAs (black
dashed line is median simulated value). Results are estimated in three
blocks, including years greater than or equal to left-hand value and
less than right-hand value.\} \label{mpa-effect}
\textbackslash end\{figure}\}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Why are the results from the channel islands important and what do
  they tell us
\item
  Put the channel islands results in context -
\end{enumerate}

Ideally then we want a control for broad environmental shocks to the
region that is independent of factors such as selection bias, biological
spillover, and fishing concentration.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  multiple lines of evidence
\item
  Theory before empirics
\item
  Patience
\end{enumerate}

The network of MPAs implemented in the Northern Channel Islands in 2003
provides an ideal opportunity to examine the regional effects of
protected area networks. This analysis is to our knowledge the first
empirical estimate of the regional effects of a large and iconic MPA
network on a broad assemblage of targeted fin-fish. Despite the well
enforced nature of these MPAs {[}xx{]}, and the presence of a rigorous
scientific monitoring program, we are unable to estimate a clear effect
of protection. While response ratios have targeted species have
increased over time, providing evidence of effective \emph{within MPA}
protection, these response ratios are on their own an unreliable
indicator of the regional conservation effects of the MPAs. Our
difference-in-difference regression estimates a high probability of
increasingly positive MPA effects over the first nine years of
protection, only to see these estimated gains reversed from 2012 to
2017. For the most recent time period, we can detect no clear positive
or negative regional effect of the Channel Island MPAs on biomass
densities of targeted fin-fish (Fig.@ref(fig:mpa-effect)).

How can we explain these results? The estimated effects from 2003 to
2014 are well in line with theoretical expectations given the species
and network in question. How might we explain the most recent results
though? One explanation may lie in fleet dynamics. Much of the
theoretical literature on MPAs assumes that all else being equal bigger
reserves produce bigger conservation gains ({\textbf{???}}). However, to
our knowledge all of these models simulate fleet dynamics through
assumptions about fishing mortality rates (e.g.~concentration of fishing
mortality (48)). The assumption of these models is that fishers
determine an amount off effort to exert, and distribute that effort
outside the MPA in response to some function.
